Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-09-01 vesicle_code/docs/permeableVes/main.tex
worked on paper a bit

We further analyze the transient dynamics by consider the vesicle
velocity~\eqref{eqn:vesVelocity} for small $\beta$. The vesicle velocity
depends on two terms: (i) the permeability velocity, $\beta(\ff \cdot
\nn)\nn$, and (ii) the force balance velocity, $\SS[\ff]$. At early
times, since $\beta \ll 1$, the permeability velocity is much smaller
than the force balance velocity and the vesicle dynamics resembles that
of a relaxing impermeable vesicle. As the force balance velocity
decreases, the permeability velocity becomes significant, and the
vesicle begins to inflate to a circle. This results in the membrane
force converging to zero \todo[inline]{Is this right?}, as does the
velocity due to permeability and force balance. Letting $\beta =
10^{-3}$, the force balance velocity (blue) and the permeability
velocity (red) are plotted at three times in
Figure~\ref{fig:vesVelocity}. We also plot the velocity of an
identically initialized impermeable vesicle (green). At $t=0.5$, the
permeability velocity is small relative to the force balance velocity,
and the force balance velocity matches with the impermeable vesicle
velocity. At time $t=5$, all three velocities are comparable in size,
and the semi-permeable vesicle begins to inflate. Then, at time $t=50$,
the impermeable vesicle is at its steady state, while the semi-permeable
vesicle is still inflating. The bottom right plot of
Figure~\ref{fig:vesVelocity} is the maximum norm of the different
velocity contributions as a function of time, and the black vertical
lines correspond to the times reported in the other three subplots.

@@

\todo[inline]{left off here}
We next investigate the interplay between the tension, flux, and
curvature of a semi-permeable vesicle. We let $\beta = 1$ and initialize
the vesicle with regions of both positive and negative curvature. The
vesicle tension, $\Lambda$, and the flux, $\ff \cdot \nn$, are plotted
in Figure~\ref{fig:starTensionFlux}. The vesicle undergoes three general
behaviors: (i) For $t \in (0,1.6 \times 10^{-2})$, the vesicle has a net
negative flux resulting in a loss of area; (ii) For $t \in (1.6 \times
10^{-2},1.0)$, the vesicle has a net positive flux and the tension along
most of the membrane is negative; (iii) For $t \in (1.0,2.0)$, the
vesicle is near its steady state and transitions to a positive constant
tension. In the first time interval, the vesicle deflates. Near the
points of largest negative curvature, there is a large negative tension
and large inflow (positive flux).  Correspondingly, near the points with
largest positive curvature, there is a small tension and outflow
(negative flux). In the second time interval, the tension is negative
and the flux is positive nearly everywhere. The regions with largest
negative curvature correspond to regions with largest negative tension
and the most inflow. Finally, in the last time interval, the tension
transitions to a positive constant function, the membrane force and flux
transition to zero, resulting in a steady state circular vesicle with
positive constant tension.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}

  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}

  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}

  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}

  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime5.pdf}
  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
  \caption{\label{fig:starTensionFlux} The vesicle tension (left) and
  flux (right) of the semi-permeable vesicle in a quiescent flow with
  $\beta=1$.}
\end{figure}

%\begin{figure}[htp]
%  \centering
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm,clip=true]{figures/StarTensionTime5.pdf}
%
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}
%  \includegraphics[width=0.18\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
%  \caption{\label{fig:starTensionFlux} The vesicle tension (top) and
%  flux (bottom) of the semi-permeable vesicle in a quiescent
%  flow with $\beta=1$.}
%\end{figure}

>>>>

We further analyze the transient dynamics by considering the vesicle
velocity~\eqref{eqn:vesVelocity} for small $\beta$. Two terms contribute
to the vesicle velocity: (i) the permeability velocity, $\beta(\ff \cdot
\nn)\nn$, and (ii) the force balance velocity, $\SS[\ff]$. At early
times, since $\beta \ll 1$, the permeability velocity is much smaller
than the force balance velocity and the vesicle dynamics resembles that
of a relaxing impermeable vesicle. As the force balance velocity
decreases, the permeability velocity becomes significant, and the
vesicle begins to inflate to a circle. This results in the membrane
force and velocity both converging to zero. Letting $\beta = 10^{-3}$,
the force balance velocity (blue) and the permeability velocity (red)
are plotted at three times in Figure~\ref{fig:vesVelocity}. We also plot
the velocity of an identically initialized impermeable vesicle (green).
At the early time $t=0.5$, the permeability velocity is small relative
to the force balance velocity, and the force balance velocity matches
with the impermeable vesicle velocity. At time intermediate time $t=5$,
all three velocities are comparable in size, and the semi-permeable
vesicle begins to inflate. Finally, at time $t=50$, the impermeable
vesicle is at its steady state, while the semi-permeable vesicle is
still inflating. The bottom right plot of Figure~\ref{fig:vesVelocity}
is the maximum norm of the different velocity contributions as a
function of time, and the black vertical lines correspond to the times
reported in the other three subplots.

@@

We next investigate the interplay between the tension, transmembrane
flux, and curvature of a semi-permeable vesicle. We let $\beta = 1$ and
initialize the vesicle with regions of both positive and negative
curvature. The vesicle tension, $\Lambda$, and the flux, $\ff \cdot
\nn$, are plotted in Figure~\ref{fig:starTensionFlux}. At early times,
regions of large negative curvature have negative tension and regions
with large positive curvature have positive tension. In addition,
regions with large curvature have a large inflow of fluid (positive
flux), while the flatter regions have an outflow of fluid. At these
early times, the net effect is vesicle deflation. As the vesicle starts
to smooth out, the same general tension and flux profiles persist, but
the net flux is positive and the vesicle inflates. At later times, the
tension starts to become positive and the vesicle continues to inflate
towards a circle. Finally, the tension transitions to a positive
constant function, the membrane force and flux approach zero, and the
vesicle shape becomes positive.

%\begin{figure*}[htp]
%  \centering
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime5.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime6.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime6.pdf}
%  \caption{\label{fig:starTensionFlux} The vesicle tension (left) and
%  flux (right) of the semi-permeable vesicle in a quiescent flow with
%  $\beta=1$.}
%\end{figure*}

\begin{figure*}[htp]
  \centering
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm,clip=true]{figures/StarTensionTime5.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm,clip=true]{figures/StarTensionTime6.pdf}

  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime6.pdf}
  \caption{\label{fig:starTensionFlux} The vesicle tension (top) and
  flux (bottom) of the semi-permeable vesicle in a quiescent
  flow with $\beta=1$.}
\end{figure*}
